---
title: "CEGE0042 Assignment - data prep"
output: html_notebook
---

### Document description:
This notebook contains code to process raw data provided for CEGE0042 final assignment.
It covers:<br>

1. Project set up

2. Area selection (Islington)

3. Data sub-selection for further analysis


### 1. Project set up:

```{r}
#Packages
library(sf)
library(sp)
library(tmap)

#Working directory definition
#getwd()

#Load data
cege_data <- "./0_source_data/UJTWorkSpace.RData"
load(cege_data)
```


### 2. Area selection (Islington)

#### Flag links with data
```{r}
#In UJT there are 256 link ids with data
all_UJT_links <- colnames(UJT)
length(all_UJT_links)

#In our shape file, there are 1402 link ids
length(LCAPShp$LCAP_ID)

#Flag which links in shape file we have data for
LCAPShp$HaveDataFlag <- 0

counter <- 0
for (master_link_value in LCAPShp$LCAP_ID) {
  counter <- counter + 1
  
  if (master_link_value %in% all_UJT_links) {
    LCAPShp$HaveDataFlag[counter] <- 1
  }
  else{
    LCAPShp$HaveDataFlag[counter] <- 0
  }
  
}

#Check that sum is equal to all links we have data for
sum(LCAPShp$HaveDataFlag)
```

#### Sub-select links (visualise data availablity)
```{r}
# Link shp file to to sf object
LCAPShp_sf <- st_as_sf(LCAPShp)

# Visualise field to show links with speed data
#plot(LCAPShp_sf["HaveDataFlag"])

# Load reference shape file for London Borough
# read in polygon shp and set crs to match 
LondonBor <- st_read('./1_shape_files/shapefiles/London_Borough_Excluding_MHW.shp') %>%
  st_transform(LondonBor, crs = st_crs(LCAPShp_sf))

# for interactive viewing 
tmap_mode("view")

# layer the boroughs, all links, and selected links 
tm_shape(LondonBor)+
  tm_polygons(col='white')+
  tm_shape(LCAPShp_sf)+
  tm_lines(col='lightgreen')+
  tm_shape(LCAPShp_sf["LCAP_ID"])+
  tm_lines(col = "LCAP_ID")

#LCAP_ID
#HaveDataFlag
```

#### Select area(s)
```{r}
# get all links that are totally within a given borough
LCAPShp_sf_sub <- LCAPShp_sf[st_within(LCAPShp_sf, 
                              LondonBor[LondonBor$NAME == "Islington",], 
                              sparse = FALSE),]

# check number of links in the subset
link_list <- LCAPShp_sf_sub$LCAP_ID

length(link_list)
# export list of links
print(link_list)
write.csv(x = link_list,"./2_data_split/selected_links.csv")
```

#### visualise links
```{r}
# plot to take a look
plot(LCAPShp_sf_sub)

# for interactive viewing 
tmap_mode("view")

# layer the boroughs, all links, and selected links 
tm_shape(LondonBor)+
  tm_polygons(col='white')+
  tm_shape(LCAPShp_sf)+
  tm_lines(col='lightgreen')+
  tm_shape(LCAPShp_sf_sub)+
  tm_lines(col='red')
```


### 3. Data sub-selection for further analysis

#### 3.1 UJT table
UJT table - each link (column) across time (row)
```{r}
# Export all source data into python friendly .csv files

# Take selected links and export selections of source data

# select relevant columns

counter = 0
link_list_string <- c()

for (entry in link_list) {
  #print(entry)
  counter <- counter + 1
  temp_string <- toString(entry)
  link_list_string[[counter]] <- temp_string
}
link_list
link_list_string

sort(colnames(UJT))

UJT[,c("2464")]


```




